\documentclass[11pt, oneside,french]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage[frenchb]{babel}
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

\title{Exercises on Strategy}
\author{20851-W2020}
\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle

\section{Demand Curves}
For the following preferences, and for consumers with exogeneous income $I$, find the infra and marginal effects of a price increase of good $X$, $p_X$ on the revenue of a monopolist:
\begin{enumerate}
    \item $U(X,Y) = (D_0 - \frac{\alpha}{2}X)X + Y$
    \item $U(X,Y) = X^{\frac{1}{3}}Y^{\frac{2}{3}}$
    \item $U(X,Y) = \min(2X,Y)$
\end{enumerate}

\section{Monopoly}

\begin{enumerate}
    \item For the quasi-linear case, find the optimal quantity produced by a monopoly and the monopoly price if marginal cost is constant. 
    \item For the quasi-linear specification of preferences above, and a constant marginal cost $c$ find the expression of the welfare loss associated with monopoly behavior. 
\end{enumerate}

\section{Duopoly I}

\begin{enumerate}
    \item For the quasi-linear case above, a new firm enters the market with a marginal cost $c_2 = \gamma c_1$.  
    \item Find the new equilibrium quantity produced by each firm if they compete on quantity (Cournot). Analyze the effect of considering $\gamma$ below, equal and above one.  
    \item Under Bertrand competition what would be the quantity produced by each firm and the price observed? Analyze the effect of considering $\gamma$ below, equal and above one.
\end{enumerate}

\section{Duopoly II}

\begin{enumerate}
    \item Consider the case with preferences represented by case (a) above. Consider the case with $D_0 = 100,\alpha=1$. Assume marginal cost of each firms are 2 and 1 respectively (firm 1 and 2). Find the expressions of the best response functions and plot in the space ($X_1,X_2$). 
    \item Find the equilibrium quantity produced by these two firms and the price. 
    \item Compare to the competitive equilibrium case. 
    \item Each of these firms are from different countries. Government of firm 1 has announced it will not subsidize the production from this firm while government of firm 2 has decided to pay 25\% of the (marginal) cost of production of firm 2. Find the new quantities produced and price observed under Cournot competition. 
\end{enumerate}


\end{document} 
