\documentclass[11pt, oneside,french]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage[frenchb]{babel}
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

\title{Exercises on Production}
\author{20851-W2020}
\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle

\title{Exercises on Production}
\author{20851-W2020}
\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle


\section{Production Functions}

Consider the production function where 

$$Y = F(X,Z) = A X^{\alpha_X} Z^{\alpha_Z} $$

The price of inputs are $p_X$ and $p_Z$ respectively. 

\begin{enumerate}
    \item Find the expression of the technical marginal rate of substitution.
    \item What are the returns to scale of this function?
    \item Find the conditional demand function for each inputs. 
    \item Find the cost function for this production function. Is the cost function convex or concave with respect to output? Relate to returns to scale
\end{enumerate}

\section{Cost Functions}

Consider the following cost function: 

$$C(p_X,p_Z,Y) = A p_X^{\alpha}p_Z^{1-\alpha} Y$$

\begin{enumerate}
    \item What are the returns to scale of the production function that yields this cost function?
    \item What are the conditional demand functions?
    \item What is the marginal cost function?
\end{enumerate}

\section{Profit Maximization}

Consider the cost function, 

$$ C(p_X,p_Z,Y) = A p_X^{\alpha}p_Z^{1-\alpha} Y^{\gamma} $$

where $\gamma>1$. The output produced by the firm is sold at price $p_Y$, find the supply function of $Y$ as a function of $(p_X,p_Z,p_Y)$. 

\section{Supply and Demand in Action}

\begin{enumerate}
    \item From the previous exercise, find the inverse supply function ($p_Y$ as a function of $(Y,p_X,p_Z)$. 
    \item If preferences of agents for good $Y$ yield the inverse demand curve $p_Y = \psi - \phi Y$, and $\gamma=1$, find the equilibrium price and quantity if the firm and consumers are price takers. Show graphically. 
    \item If the price of the input $p_X$ increases, how does the equilibrium price and quantity of $Y$ changes. 
\end{enumerate}


 

\end{document} 
\end{document}  